---
output: 
  stevetemplates::article:
    fig_caption: true
#bibliography: master.bib
biblio-style: apsr
title: "GAM calibration report"
# thanks: "Replication files are available on the author's Github account (http://github.com/svmiller). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: svmille@clemson.edu."
author:
- name: Victor Van der Meersch
  affiliation: CEFE, CNRS
abstract: "This document provides detailed information about a specific correlative SDM. "
# keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
#geometry: margin=1cm
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
endnote: no
header-includes:
  - \usepackage{xcolor}
  - \usepackage{graphicx}
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage[width=.6\textwidth, textfont=it, font=small]{caption}
  - \usepackage{nopageno}
  - \setlength{\voffset}{-0.4in}
  - \geometry{top=1in, bottom=-1in, left=1.25in, right=1.25in}

---

```{r setup, include=FALSE}
wd <- "C:/Users/vandermeersch/Documents/CEFE/phd/correlative_models"

source("C:/Users/vandermeersch/Documents/CEFE/phd/correlative_models/valavi_et_al/ecm1486-sup-0003-datas1/DataS1/modelling_codes/prediction_helper.R")

library(AUC)
library(ggplot2)
library(colorspace)
library(dplyr)
library(sf)
library(raster)
library(rnaturalearth)
library(ecospat)
library(abind)
library(Matrix)
library(glmnet)
library(dismo)
library(mgcv)
library(disdat)

knitr::opts_chunk$set(cache=TRUE,
                      message=FALSE, warning=FALSE,
                      fig.path='figs/',
                      cache.path = '_cache/',
                      fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      })


```

```{r data, include=FALSE}
# model folder
model_type <- "maxent"
mod_dir <- file.path(wd, model_type)
fit_file <- file.path(mod_dir, "fit", "fagus_sylvatica", "maxent_defpred_2022-06-14.rds")
outfile <- readRDS(fit_file)

# load all species presence/absence
sp_data_dir <- "D:/species/processed"
load(file.path(sp_data_dir, "fagus_sylvatica", "fagus_sylvatica_presabs.Rdata"))

# # Covars
# clim_data_dir <- "D:/climate/ERA5-Land/bioclim_format"
# soil_data_dir <- "D:/soil/processed"
# ## bioclim data from ERA5-Land
# biovars_all <- list()
# for(i in as.character(1970:2000)){
#   load(paste0(clim_data_dir, "/biovars_", i, ".Rdata"))
#   biovars_all <- append(biovars_all, list(biovars))
# }
# rm(biovars)
# biovars_30y <- abind(biovars_all, along=3)
# biovars_30y <- apply(biovars_30y, c(1,2), mean)
# biovars_30y <- as.data.frame(biovars_30y)
# biovars_30y$lat <- round(biovars_30y$lat,1)
# biovars_30y$lon <- round(biovars_30y$lon,1)
# ## soil data
# load(file.path(soil_data_dir, "data_soil.Rdata"))


# predictors
wd_pred <- "C:/Users/vandermeersch/Documents/CEFE/phd/correlative_models/predictors"
load(file.path(wd_pred, "predictors_data.Rdata"))

europe_data <- predictors_data %>% 
  dplyr::select(c("lon", "lat", all_of(outfile$covars)))

## for pres/abs points
presabs_data <- left_join(fagussylvatica_presabs, predictors_data, by = c("lon", "lat")) %>% 
  dplyr::select(c("lon", "lat", "pres", all_of(outfile$covars)))




```

```{r gam_pred, eval = (model_type == "gam"), include=FALSE, eval=F}

# Predictions with GAM
europe_data$fitness <- outfile$europe_prediction
presabs_data$fitness <- as.numeric(predict(outfile$model, presabs_data, type = "response"))

```

```{r lasso_pred, eval = (model_type == "lasso_glm"), include=FALSE, eval=F}


# Predictions with Lasso GLM
quad_obj <- make_quadratic(europe_data, cols = outfile$covars)
europe_quad <- predict.make_quadratic(quad_obj, newdata = europe_data)
new_vars <- names(europe_quad)[!(names(europe_quad) %in% c("pres", "lon", "lat"))]
europe_sparse <- sparse.model.matrix(~. -1, europe_quad[, new_vars])
europe_data$fitness <- as.numeric(predict(outfile$model, europe_sparse, type = "response", s = "lambda.min"))
quad_obj <- make_quadratic(presabs_data, cols = outfile$covars)
presabs_quad <- predict.make_quadratic(quad_obj, newdata = presabs_data)
presabs_sparse <- sparse.model.matrix(~. -1, presabs_quad[, new_vars])
presabs_data$fitness <- as.numeric(predict(outfile$model, presabs_sparse, type = "response", s = "lambda.min"))

```

```{r maxent_pred, eval = (model_type == "maxent"), include=FALSE, eval=T}

# Predictions with Maxent
europe_data$fitness <- predict(outfile$model, europe_data, args = c("outputformat=cloglog")) # all Europe
presabs_data$fitness <- predict(outfile$model, presabs_data, args = c("outputformat=cloglog")) # only on pres/abs points

```

```{r calibration_points, include=FALSE}
# Only on calibration points : below...
# # Simulation only on calibration points
# calsim_dir <- file.path(sim_dir, "calibration_points")
# fitness_cal <- read_mean_outputvalue(calsim_dir, "Fitness")
# 
# # Best threshold maximizing Youden index
# youden_index <- sensitivity(fitness_cal, factor(species_occurrence$pres))$measure +
#   specificity(fitness_cal, factor(species_occurrence$pres))$measure -1
# thresholds <- roc(fitness_cal, factor(species_occurrence$pres))$cutoffs
# best_threshold <- thresholds[which(youden_index == max(youden_index))]

# On every points
youden_index <- sensitivity(presabs_data$fitness, as.factor(presabs_data$pres))$measure +
  specificity(presabs_data$fitness, as.factor(presabs_data$pres))$measure -1
thresholds <- roc(presabs_data$fitness, as.factor(presabs_data$pres))$cutoffs
best_threshold <- thresholds[which(youden_index == max(youden_index))]






```

## 1. Insights on the calibration process

Input data:  
- Species : *F. sylvatica*   
- `r nrow(outfile$occ_data[outfile$occ_data$pres == 1, ])` presence points and `r nrow(outfile$occ_data[outfile$occ_data$pres == 0, ])` background points  
\
Calibration performance:  
- Runtime = `r outfile$runtime` seconds  
- AUC on calibration points = `r round(outfile$auc_cal,3)` (warning: not significant, background points...)  
- Discrimination metric: AUC on all presence/absence points = `r round(outfile$auc_tot,3)`  

```{r calpoints_map, echo=FALSE, fig.height = 5, fig.width = 5, fig.align = "center", fig.cap = "Calibration points: presences points in green and background points in orange. Light green is species distribution area.", fig.pos="H"}

# Europe map
countries_EUForest <- c("Austria", "Belarus", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czech Rep.", "Denmark", "Estonia", "Finland", "France", "Germany",
                        "Greece", "Hungary", "Ireland", "Italy", "Latvia", "Lithuania", "Moldova", "Netherlands", "Norway", "Poland", "Portugal", "Romania",
                        "Slovakia", "Slovenia", "Spain", "Sweden", "Switzerland", "United Kingdom")
other_countries <- c("Ukraine", "Bosnia and Herzegovina", "Republic of Serbia", "Macedonia", "Greece",
                     "Kosovo", "Albania", "Montenegro")
countries <- c(countries_EUForest, other_countries)
world_map <- ne_countries(scale="medium",returnclass = 'sf')
eu_map <- world_map %>% filter(sovereignt %in% countries)
eu_map_cropped <- eu_map %>% 
  st_crop(st_bbox(c(xmin = -12, xmax = 40, ymax = 71, ymin = 34), crs = st_crs(4326)))


ggplot() + 
  geom_sf(data = eu_map_cropped, fill = 'grey', alpha=0.1, color='grey', size = 0.1) +
  geom_tile(data = outfile$occ_data[outfile$occ_data$pres == 0, ], aes(x = lon, y = lat), fill = "#D57A66", alpha=0.2) +
  geom_tile(data= fagussylvatica_presabs[fagussylvatica_presabs$pres == 1,], aes(x = lon, y = lat), fill = "#8FD5A6", alpha=0.3) +
  geom_tile(data = outfile$occ_data[outfile$occ_data$pres == 1, ], aes(x = lon, y = lat), fill = "#0C8346", alpha=1) +
  theme_void() +
  theme(legend.position="none") +
  theme(plot.title = element_text(size=12))



```


## 2. Simulation at European scale


```{r fitness_map, echo=FALSE, fig.height = 4.3, fig.width = 4.3, fig.align = "center", fig.cap = "Fitness map with calibrated parameter set", fig.pos="H"}

ggplot(data=europe_data, aes(x = lon, y = lat)) + 
  geom_tile(aes(fill = fitness), color = NA) +
  theme_void() +
  theme(legend.position="bottom") +
  theme(legend.title=element_blank()) +
  ylab("") +
  xlab("") +
  scale_fill_gradient(low = "#EBEBD3", high = "#488B49", limits=c(0,1)) +
  theme(legend.key.height= unit(0.2, 'cm'),
        legend.key.width= unit(1, 'cm'))

```



```{r presence_map, echo=FALSE, fig.height = 4.3, fig.width = 4.3, fig.align = "center", fig.cap = paste0("Species presence map with calibrated parameter set. AUC threshold maximizing Youden index = ", round(best_threshold,3), "."), fig.pos="H", eval=T}
data_presence <- europe_data
data_presence[data_presence$fitness < best_threshold, 'presence'] <- 0
data_presence[data_presence$fitness >= best_threshold, 'presence'] <- 1

ggplot(data=data_presence, aes(x = lon, y = lat)) + 
  geom_raster(aes(fill = presence), color = NA) +
  theme_void() +
  theme(legend.position="none") +
  ylab("") +
  xlab("") +
  scale_fill_gradient(low = "#EBEBD3", high = "#488B49")

```

<!--
# References
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\noindent
-->
