---
output: 
  stevetemplates::article:
    fig_caption: true
#bibliography: master.bib
biblio-style: apsr
title: "Phenofit calibration report"
# thanks: "Replication files are available on the author's Github account (http://github.com/svmiller). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: svmille@clemson.edu."
author:
- name: Victor Van der Meersch
  affiliation: CEFE, CNRS
abstract: "This document provides detailed information about a specific calibration run, including CMA-ES algorithm performance and model outputs. "
# keywords: "pandoc, r markdown, knitr"
date: "`r format(Sys.time(), '%B %d, %Y')`"
#geometry: margin=1cm
fontfamily: mathpazo
fontsize: 11pt
# spacing: double
endnote: no
header-includes:
  - \usepackage{graphicx}
  - \usepackage{float}
  - \floatplacement{figure}{H}
  - \usepackage[width=.6\textwidth, textfont=it, font=small]{caption}
  - \usepackage{nopageno}
  - \setlength{\voffset}{-0.4in}
  - \geometry{top=1in, bottom=-1in, left=1.25in, right=1.25in}

---

```{r setup, include=FALSE}
wd <- "C:/Users/vandermeersch/Documents/CEFE/thesis/phenofit/calibration_report"

library(AUC)
library(ggplot2)
library(colorspace)
library(dplyr)
library(sf)
library(raster)
library(rnaturalearth)

knitr::opts_chunk$set(cache=TRUE,
                      message=FALSE, warning=FALSE,
                      fig.path='figs/',
                      cache.path = '_cache/',
                      fig.process = function(x) {
                      x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
                      if (file.rename(x, x2)) x2 else x
                      })

source(file.path(wd, "functions", "read_mean_outputvalue.R"))

# Way to get lat/lon
climate_folder <- "D:/climate/ERA5-Land/phenofit_format/transformed"
alt_file <- paste0(climate_folder, "/ERA5LAND_", "Altitude.fit")
alt <- fread(alt_file, showProgress=F)
colnames(alt) <- c("lat", "lon", "alt")
alt$lat <- round(alt$lat, 1)
alt$lon <- round(alt$lon, 1)

```

```{r data, include=FALSE}
# Calibration and simulation (at European scale) folders
cal_dir <- "D:/calibrations/phenofit/abies_alba/1000pres_1000abs/model9"
sim_dir <- "D:/simulations/phenofit/backward/abies_alba/model9"

# Calibration points
cal_points <- "D:/species/processed/abies_alba/1000pres_1000abs/occurrence_subset_3.Rdata"
load(cal_points)

# Species presence points and pres/abs points
load("D:/species/processed/abies_alba/abiesalba_occ_rs.Rdata")
load("D:/species/processed/abies_alba/abiesalba_presabs.Rdata")

cmaes_fit <- list.files(path = cal_dir, pattern = "\\.Rdata$", full.names = T)
load(cmaes_fit)

# Load fitness
fitness <- read_mean_outputvalue(sim_dir, "Fitness") # all Europe
fitness_presabs <- read_mean_outputvalue(sim_dir, "Fitness", points = abiesalba_presabs) # only on pres/abs points

# AUC on pres/abs points 
auc_presabs <- auc(roc(fitness_presabs, as.factor(abiesalba_presabs$pres)))

```

```{r calibration_points, include=FALSE}
# Only on calibration points : below...
# # Simulation only on calibration points
# calsim_dir <- file.path(sim_dir, "calibration_points")
# fitness_cal <- read_mean_outputvalue(calsim_dir, "Fitness")
# 
# # Best threshold maximizing Youden index
# youden_index <- sensitivity(fitness_cal, factor(species_occurrence$pres))$measure +
#   specificity(fitness_cal, factor(species_occurrence$pres))$measure -1
# thresholds <- roc(fitness_cal, factor(species_occurrence$pres))$cutoffs
# best_threshold <- thresholds[which(youden_index == max(youden_index))]

# On every points
youden_index <- sensitivity(fitness_presabs, as.factor(abiesalba_presabs$pres))$measure +
  specificity(fitness_presabs, as.factor(abiesalba_presabs$pres))$measure -1
thresholds <- roc(fitness_presabs, as.factor(abiesalba_presabs$pres))$cutoffs
best_threshold <- thresholds[which(youden_index == max(youden_index))]






```

## 1. Insights on the calibration process

Input data:  
- Species : *A. alba*   
- `r nrow(species_occurrence)` calibration points (50/50 presence/absence)  
\
Parameter tuning and calibration performance:  
- $\lambda = 20$, $\mu = 10$, $\sigma = 2$   
- $20$ parallel objective function evaluations  
- $60$ CPUs, $90 Go$ of memory  
- Runtime = `r round(cmaes_fit[[1]][["runtime"]]/3600, 2)` hours  
- AUC on calibration points = `r 1-round(cmaes_fit[[1]][["value"]], 3)`  
- AUC on all presence/absence points = `r round(auc_presabs, 3)`

```{r calpoints_map, echo=FALSE, fig.height = 5, fig.width = 5, fig.align = "center", fig.cap = "Calibration points: presence in red and absence in orange. Light green is species distribution area.", fig.pos="H"}

# Europe map
countries_EUForest <- c("Austria", "Belarus", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czech Rep.", "Denmark", "Estonia", "Finland", "France", "Germany",
                        "Greece", "Hungary", "Ireland", "Italy", "Latvia", "Lithuania", "Moldova", "Netherlands", "Norway", "Poland", "Portugal", "Romania",
                        "Slovakia", "Slovenia", "Spain", "Sweden", "Switzerland", "United Kingdom")
other_countries <- c("Ukraine", "Bosnia and Herzegovina", "Republic of Serbia", "Macedonia", "Greece",
                     "Kosovo", "Albania", "Montenegro")
countries <- c(countries_EUForest, other_countries)
world_map <- ne_countries(scale="medium",returnclass = 'sf')
eu_map <- world_map %>% filter(sovereignt %in% countries)
eu_map_cropped <- eu_map %>% 
  st_crop(st_bbox(c(xmin = -12, xmax = 40, ymax = 71, ymin = 34), crs = st_crs(4326)))

abiesalba_occ_df <- na.omit(as.data.frame(abiesalba_occ_rs, xy=T))

ggplot() + 
  geom_sf(data = eu_map_cropped, fill = 'grey', alpha=0.1, color='grey', size = 0.1) +
  geom_tile(data=abiesalba_occ_df, aes(x, y), fill = "#8FD5A6", alpha=0.3) +
  geom_tile(data = species_occurrence[species_occurrence$pres==1,], aes(x = lon, y = lat), fill = "#0C8346", alpha=1) +
  geom_tile(data = species_occurrence[species_occurrence$pres==0,], aes(x = lon, y = lat), fill = "#D57A66", alpha=0.7) +
  theme_void() +
  theme(legend.position="none") +
  theme(plot.title = element_text(size=12))



```


## 2. Simulation at European scale


```{r fitness_map, echo=FALSE, fig.height = 4.3, fig.width = 4.3, fig.align = "center", fig.cap = "Fitness map with calibrated parameter set", fig.pos="H"}
data_fitness <- as.data.frame(cbind(alt$lat, alt$lon,t(fitness)))
colnames(data_fitness) <- c("lat", "lon", "fitness")

ggplot(data=data_fitness, aes(x = lon, y = lat)) + 
  geom_tile(aes(fill = fitness), color = NA) +
  theme_void() +
  theme(legend.position="bottom") +
  theme(legend.title=element_blank()) +
  ylab("") +
  xlab("") +
  scale_fill_gradient(low = "#EBEBD3", high = "#488B49", limits=c(0,1)) +
  theme(legend.key.height= unit(0.2, 'cm'),
        legend.key.width= unit(1, 'cm'))

```



```{r presence_map, echo=FALSE, fig.height = 4.3, fig.width = 4.3, fig.align = "center", fig.cap = paste0("Species presence map with calibrated parameter set. AUC threshold maximizing Youden index = ", round(best_threshold,3), "."), fig.pos="H"}
data_presence <- data_fitness
data_presence[data_presence$fitness < best_threshold, 'presence'] <- 0
data_presence[data_presence$fitness >= best_threshold, 'presence'] <- 1

ggplot(data=data_presence, aes(x = lon, y = lat)) + 
  geom_tile(aes(fill = presence), color = NA) +
  theme_void() +
  theme(legend.position="none") +
  ylab("") +
  xlab("") +
  scale_fill_gradient(low = "#EBEBD3", high = "#488B49")

```

<!--
# References
\setlength{\parindent}{-0.2in}
\setlength{\leftskip}{0.2in}
\setlength{\parskip}{8pt}
\vspace*{-0.2in}
\noindent
-->
